SPP-Perc: Peceptron Learning Based Enhanced Signature Path Prefetcher

### 2018-11-17 design.tex
Aligning Table 3

\item \textbf{Program Counter XOR Depth}: PC is for the instruction
  which triggered the prefetch chain.  Depth refers to the iteration
  count of the lookahead stages.  In general, PC is not considered as
  a good basis for doing lookahead prefetching as all the prefetches
  with depth >= 1 gets aliased onto the same PC which won't be true in
  actual demand fetch case.  This feature resolves a PC into a
  different value for each lookahead depth of prefetch speculation, giving
  a more accurate correlation in lookahead cases.  This is akin to the
  concept of Virtual Program Counters\cite{VPC} introduced by Kim
  \textit{et. al.} for indirect branch prediction.

@@

\item \textbf{Program Counter XOR Delta}: This tells us if a given PC
  favours particular value(s) of delta.  As noted earlier, while PC in
  itself does not much useful information, this hash resolves PC into
  different values based on the tendency of that PC to favour a
  certain delta.  This way, dynamic nature of different instances of
  the same memory access instruction can be captured here.

@@

\item \textbf{Feature Selection:} Six out of the nine features we developed use information derived 
from program execution, irrespective of the baseline SPP prefetcher. Beyond that, the feature set 
can be expanded to convey any useful information from the prefetcher to the perceptron throttler.
For example, in our implementation 'Confidence', 'Current Signature XOR Delta', and 
'Page Address XOR Confidence' are such features.

>>>>

\item \textbf{Program Counter XOR Depth}: PC is for the instruction
  which triggered the prefetch chain.  Depth refers to the iteration
  count of the lookahead stages.  In general, PC is not considered as
  a good basis for doing lookahead prefetching as all the prefetches
  with depth >= 1 gets aliased onto the same PC which won't be true in
  actual demand fetch case.  This feature resolves a PC into a
  different value for each lookahead depth of prefetch speculation, giving
  a more accurate correlation in lookahead cases.  This is akin to the
  concept of Virtual Program Counters~\cite{VPC} introduced by Kim
  \textit{et. al.} for indirect branch prediction.

@@

\item \textbf{Program Counter XOR Delta}: This tells us if a given PC
  favours particular value(s) of delta.  As noted earlier, while PC in
  itself does not convey much useful information, this hash resolves PC into
  different values based on the tendency of that PC to favour a
  certain delta.  This way, dynamic nature of different instances of
  the same memory access instruction can be captured here.

@@

\item \textbf{Feature Selection:} Six out of the nine features we developed use information derived 
from program execution, irrespective of the baseline SPP prefetcher. Beyond that, the feature set 
can be expanded to convey any useful information from the prefetcher to the perceptron throttler.
For example, in our implementation `Confidence', `Current Signature XOR Delta', and 
'Page Address XOR Confidence' are such features.

### 2018-11-17 intro.tex
Aligning Table 3

\section{INTRODUCTION}
\label{Introduction}
Since past few decades, processor and memory scaling has been enabled
with different philosophies in mind.  While processor scaling focused
on speed improvements, memory scaling was centred around increasing
the storage capacity.  The process technology reaching scaling
limitations led to Memory Wall\cite{MemWall} - the increasing
performance gap between the processor speed and the memory speeds.
Various techniques have been developed to combat this widening gap.
Data prefetching is one such important technique.  Data prefetching
exploits the fact that in most applications, memory accesses have a
well-defined pattern.

@@

\begin{itemize}
\item This is the first time where a truly on-line learning based
  neural network model has been used when it comes to hardware data
  prefetching.  The prior works in this area either relied on program
  semantics\cite{Semantics} or were application
  specific\cite{Datacenter} or could not yield a significant
  improvement over the baseline prefetcher used\cite{BadPerc}.

@@

\item Examining the perceptron weights doing a statistical study to
  quantify the contribution made by each feature.  We have also used
  this idea to develop a comprehensive set of features where each
  feature contributes some `new' information to the perceptron.  The
  details are explained under methodology, in Section \ref{Method-Overheads}.

@@

We tested NATCG on Champsim, an open source simulator which models
an Out-of-Order CPU using.  For evaluation, we used SPEC 2017
benchmarks.  On a single core configuration, NATCH increases
performance of the system by 39.5\% as compared to the baseline with
no prefetching, on the memory intensive subset of the benchmark.  This
outperforms the next best prefetcher, SPP, by 6.84\%.  On a multi-core
system running a mix of memory intensive subset of the SPEC 2017
traces, NATCH saw an improvement of 11.9\% over SPP for 4-core
system and 10.67\% for 8-core system.

>>>>

\section{INTRODUCTION}
\label{Introduction}
Since past few decades, processor and memory scaling has been enabled
with different philosophies in mind.  While processor scaling focused
on speed improvements, memory scaling was centred around increasing
the storage capacity.  The process technology reaching scaling
limitations led to Memory Wall~\cite{MemWall} - the increasing
performance gap between the processor speed and the memory speeds.
Various techniques have been developed to combat this widening gap.
Data prefetching is one such important technique.  Data prefetching
exploits the fact that in most applications, memory accesses have a
well-defined pattern.

@@

\begin{itemize}
\item This is the first time where a truly on-line learning based
  neural network model has been used when it comes to hardware data
  prefetching.  The prior works in this area either relied on program
  semantics~\cite{Semantics} or were application
  specific~\cite{Datacenter} or could not yield a significant
  improvement over the baseline prefetcher used~\cite{BadPerc}.

@@

\item Examining the perceptron weights by doing a statistical study to
  quantify the contribution made by each feature.  We have also used
  this idea to develop a comprehensive set of features where each
  feature contributes some `new' information to the perceptron.  The
  details are explained under methodology, in Section \ref{Method-Overheads}.

@@

We tested NATCH on Champsim, an open source simulator which models
an Out-of-Order CPU using.  For evaluation, we used SPEC 2017
benchmarks.  On a single core configuration, NATCH increases
performance of the system by 39.5\% as compared to the baseline with
no prefetching, on the memory intensive subset of the benchmark.  This
outperforms the next best prefetcher, SPP, by 6.84\%.  On a multi-core
system running a mix of memory intensive subset of the SPEC 2017
traces, NATCH saw an improvement of 11.9\% over SPP for 4-core
system and 10.67\% for 8-core system.
